\newpage
\chapter{Nagavagga: The Elephant}
320. As an elephant in the battlefield withstands arrows shot from bows all around, even so shall I endure abuse. There are many, indeed, who lack virtue.

321. A tamed elephant is led into a crowd, and the king mounts a tamed elephant. Best among men is the subdued one who endures abuse.

322. Excellent are well-trained mules, thoroughbred Sindhu horses and noble tusker elephants. But better still is the man who has subdued himself.

323. Not by these mounts, however, would one go to the Untrodden Land (Nibbana), as one who is self-tamed goes by his own tamed and well-controlled mind.

324. Musty during rut, the tusker named Dhanapalaka is uncontrollable. Held in captivity, the tusker does not touch a morsel, but only longingly calls to mind the elephant forest.

325. When a man is sluggish and gluttonous, sleeping and rolling around in bed like a fat domestic pig, that sluggard undergoes rebirth again and again.

326. Formerly this mind wandered about as it liked, where it wished and according to its pleasure, but now I shall thoroughly master it with wisdom as a mahout controls with his ankus an elephant in rut.

327. Delight in heedfulness! Guard well your thoughts! Draw yourself out of this bog of evil, even as an elephant draws himself out of the mud.

328. If for company you find a wise and prudent friend who leads a good life, you should, overcoming all impediments, keep his company joyously and mindfully.

329. If for company you cannot find a wise and prudent friend who leads a good life, then, like a king who leaves behind a conquered kingdom, or like a lone elephant in the elephant forest, you should go your way alone.

330. Better it is to live alone; there is no fellowship with a fool. Live alone and do no evil; be carefree like an elephant in the elephant forest.

331. Good are friends when need arises; good is contentment with just what one has; good is merit when life is at an end, and good is the abandoning of all suffering (through Arahantship).

332. In this world, good it is to serve one's mother, good it is to serve one's father, good it is to serve the monks, and good it is to serve the holy men.

333. Good is virtue until life's end, good is faith that is steadfast, good is the acquisition of wisdom, and good is the avoidance of evil.
