\chapter{Cittavagga: The Mind}
\epigraphhead[0]{\epigraph{\textit{``Believe nothing, no matter where you read it, or who said it, no matter if I have said it, unless it agrees with your own reason and your own common sense."}}{Gautama Buddha}}
33. Just as a fletcher straightens an arrow shaft, even so the discerning man straightens his mind — so fickle and unsteady, so difficult to guard.

34. As a fish when pulled out of water and cast on land throbs and quivers, even so is this mind agitated. Hence should one abandon the realm of \Gls{Mara}.

35. Wonderful, indeed, it is to subdue the mind, so difficult to subdue, ever swift, and seizing whatever it desires. A tamed mind brings happiness.

36. Let the discerning man guard the mind, so difficult to detect and extremely subtle, seizing whatever it desires. A guarded mind brings happiness.

37. Dwelling in the cave (of the heart), the mind, without form, wanders far and alone. Those who subdue this mind are liberated from the bonds of \Gls{Mara}.

38. Wisdom never becomes perfect in one whose mind is not steadfast, who knows not the Good Teaching and whose faith wavers.

39. There is no fear for an awakened one, whose mind is not sodden (by lust) nor afflicted (by hate), and who has gone beyond both merit and demerit.\footnote{The arahant is said to be beyond both merit and demerit because, as he has abandoned all defilements, he can no longer perform evil actions; and as he has no more attachment, his virtuous actions no longer bear kammic fruit.}

40. Realizing that this body is as fragile as a clay pot, and fortifying this mind like a well-fortified city, fight out \Gls{Mara} with the sword of wisdom. Then, guarding the conquest, remain unattached.

41. Ere long, alas! this body will lie upon the earth, unheeded and lifeless, like a useless log.

42. Whatever harm an enemy may do to an enemy, or a hater to a hater, an ill-directed mind inflicts on oneself a greater harm.

43. Neither mother, father, nor any other relative can do one greater good than one's own well-directed mind.
