\newpage
\chapter{Sahassavagga: The Thousands}

100. Better than a thousand useless words is one useful word, hearing which one attains peace.

101. Better than a thousand useless verses is one useful verse, hearing which one attains peace.

102. Better than reciting a hundred meaningless verses is the reciting of one verse of Dhamma, hearing which one attains peace.

103. Though one may conquer a thousand times a thousand men in battle, yet he indeed is the noblest victor who conquers himself.

104-105. Self-conquest is far better than the conquest of others. Not even a god, an angel, Mara or Brahma\footnote{Brahma: a high divinity in ancient Indian religion.} can turn into defeat the victory of a person who is self-subdued and ever restrained in conduct.

106. Though month after month for a hundred years one should offer sacrifices by the thousands, yet if only for a moment one should worship those of perfected minds that honor is indeed better than a century of sacrifice.

107. Though for a hundred years one should tend the sacrificial fire in the forest, yet if only for a moment one should worship those of perfected minds, that worship is indeed better than a century of sacrifice.

108. Whatever gifts and oblations one seeking merit might offer in this world for a whole year, all that is not worth one fourth of the merit gained by revering the Upright Ones, which is truly excellent.

109. To one ever eager to revere and serve the elders, these four blessing accrue: long life and beauty, happiness and power.

110. Better it is to live one day virtuous and meditative than to live a hundred years immoral and uncontrolled.

111. Better it is to live one day wise and meditative than to live a hundred years foolish and uncontrolled.

112. Better it is to live one day strenuous and resolute than to live a hundred years sluggish and dissipated.

113. Better it is to live one day seeing the rise and fall of things than to live a hundred years without ever seeing the rise and fall of things.

114. Better it is to live one day seeing the Deathless than to live a hundred years without ever seeing the Deathless.

115. Better it is to live one day seeing the Supreme Truth than to live a hundred years without ever seeing the Supreme Truth.
