\newglossaryentry{Mara}
{
  name=Mara,
  description={the Tempter in Buddhism, represented in the scriptures as an evil-minded deity who tries to lead people from the path to liberation. The commentaries explain Mara as the lord of evil forces, as mental defilements and as death}
}

\newglossaryentry{}
{
  name=Âgîvika,
 description=one belonging to a sect of naked ascetics
}

\newglossaryentry{}
{
  name=Arahat,
 description=a holy man, a saint
}

\newglossaryentry{}
{
  name=Ariya,
 description=noble
}

\newglossaryentry{}
{
  name=Bhagavat,
 description=worshipful; blessed, a name of a Buddha
}

\newglossaryentry{}
{
  name=Bhikkhu,
 description=a mendicant
}

\newglossaryentry{}
{
  name=Brahman,
 description=the supreme god of the Hindus
}

\newglossaryentry{}
{
  name=Brâhmana,
 description=a sage
}

\newglossaryentry{}
{
  name=Buddha,
 description=enlightened, a name of certain holy men who have freed themselves from existence, particularly of Samana Gotama
}

\newglossaryentry{}
{
  name=Dhamma,
 description=tenet, doctrine, custom, law, religion, virtue, thing
}

\newglossaryentry{}
{
  name=Gahattha,
 description=Gihin, a householder
}

\newglossaryentry{}
{
  name=Gotama,
 description=a name of the last Buddha
}

\newglossaryentry{}
{
  name=Gatila,
 description=an ascetic wearing clotted hair
}

\newglossaryentry{}
{
  name=Gina,
 description=a conqueror, a name of a Buddha
}

\newglossaryentry{}
{
  name=Isi,
 description=a sage
}

\newglossaryentry{}
{
  name=Khattiya,
 description=a warrior, a prince
}

\newglossaryentry{}
{
  name=Kandâla,
 description=an outcast
}

\newglossaryentry{}
{
  name=Mâra,
 description=a name of the king of death, the devil
}

\newglossaryentry{}
{
  name=Muni,
 description=a thinker, a sage
}

\newglossaryentry{}
{
  name=Nâga,
 description=an eminent man; sinless?
}

\newglossaryentry{}
{
  name=Namuki,
 description=another meaning of Mâra
}

\newglossaryentry{}
{
  name=Nibbâna,
 description=extinction, the state of bliss of the Buddhist
}

\newglossaryentry{}
{
  name=Nigantha,
 description=a naked ascetic
}

\newglossaryentry{}
{
  name=Pabbaggâ,
 description=leaving the world, embracing ascetic life, taking the robe
}
,
	


\newglossaryentry{}
{
  name=Pabbagita,
 description=an ascetic, having taken the robe
}

\newglossaryentry{}
{
  name=Paribbâga,
 description=Paribbâgaka, a wandering mendicant
}

\newglossaryentry{}
{
  name=,
 description=
}
Sakka = Sakya,
	

belonging to the Sakya tribe.
\newglossaryentry{}
{
  name=,
 description=
}
Sakyamuni,
	

the Sakya sage, a name of Buddha.
\newglossaryentry{}
{
  name=,
 description=
}
Samana,
	

an ascetic.
\newglossaryentry{}
{
  name=,
 description=
}
Samkhârâ,
	

all compound things, the material world.
\newglossaryentry{}
{
  name=,
 description=
}
Samsâra,
	

revolution, transmigration.
\newglossaryentry{}
{
  name=,
 description=
}
Sâvaka,
	

a hearer, a follower, a disciple of Buddha, including both laity and clergy.
\newglossaryentry{}
{
  name=,
 description=
}
Sekha,
	

a novice, student.
\newglossaryentry{}
{
  name=,
 description=
}
Sudda,
	

a man of the servile caste.
\newglossaryentry{}
{
  name=,
 description=
}
Sugata,
	

happy, a name of a Buddha.
\newglossaryentry{}
{
  name=,
 description=
}
Thera,
	

an elder, a senior priest.
\newglossaryentry{}
{
  name=,
 description=
}
Titthiya,
	

an ascetic adhering to a certain system of philosophy.
\newglossaryentry{}
{
  name=,
 description=
}
Upadhi,
	

the elements of the world.
\newglossaryentry{}
{
  name=,
 description=
}
Upâsaka,
	

a follower, a lay devotee.
\newglossaryentry{}
{
  name=,
 description=
}
Upasampadâ,
	

priest's orders.
\newglossaryentry{}
{
  name=,
 description=
}
Vessa, Vessika,
	

a man of the third caste.
\newglossaryentry{}
{
  name=,
 description=
}
Yakkha,
	

a giant, a malignant spirit.