\newpage
\chapter{Lokavagga: The World}
167. Follow not the vulgar way; live not in heedlessness; hold not false views; linger not long in worldly existence.

168. Arise! Do not be heedless! Lead a righteous life. The righteous live happily both in this world and the next.

169. Lead a righteous life; lead not a base life. The righteous live happily both in this world and the next.

170. One who looks upon the world as a bubble and a mirage, him the King of Death sees not.

171. Come! Behold this world, which is like a decorated royal chariot. Here fools flounder, but the wise have no attachment to it.

172. He who having been heedless is heedless no more, illuminates this world like the moon freed from clouds.

173. He, who by good deeds covers the evil he has done, illuminates this world like the moon freed from clouds.

174. Blind is the world; here only a few possess insight. Only a few, like birds escaping from the net, go to realms of bliss.

175. Swans fly on the path of the sun; men pass through the air by psychic powers; the wise are led away from the world after vanquishing Mara and his host.

176. For a liar who has violated the one law (of truthfulness) who holds in scorn the hereafter, there is no evil that he cannot do.

177. Truly, misers fare not to heavenly realms; nor, indeed, do fools praise generosity. But the wise man rejoices in giving, and by that alone does he become happy hereafter.

178. Better than sole sovereignty over the earth, better than going to heaven, better even than lordship over all the worlds is the supramundane Fruition of Stream Entrance\footnote{Stream-entry (sotapatti): the first stage of supramundane attainment.}.
