\newpage
\chapter{Kodhavagga: Anger}
221. One should give up anger, renounce pride, and overcome all fetters. Suffering never befalls him who clings not to mind and body and is detached.

222. He who checks rising anger as a charioteer checks a rolling chariot, him I call a true charioteer. Others only hold the reins.

223. Overcome the angry by non-anger; overcome the wicked by goodness; overcome the miser by generosity; overcome the liar by truth.

224. Speak the truth; yield not to anger; when asked, give even if you only have a little. By these three means can one reach the presence of the gods.

225. Those sages who are inoffensive and ever restrained in body, go to the Deathless State, where, having gone, they grieve no more.

226. Those who are ever vigilant, who discipline themselves day and night, and are ever intent upon Nibbana — their defilements fade away.

227. O Atula! Indeed, this is an ancient practice, not one only of today: they blame those who remain silent, they blame those who speak much, they blame those who speak in moderation. There is none in the world who is not blamed.

228. There never was, there never will be, nor is there now, a person who is wholly blamed or wholly praised.

229. But the man whom the wise praise, after observing him day after day, is one of flawless character, wise, and endowed with knowledge and virtue.

230. Who can blame such a one, as worthy as a coin of refined gold? Even the gods praise him; by Brahma, too, is he praised.

231. Let a man guard himself against irritability in bodily action; let him be controlled in deed. Abandoning bodily misconduct, let him practice good conduct in deed.

232. Let a man guard himself against irritability in speech; let him be controlled in speech. Abandoning verbal misconduct, let him practice good conduct in speech.

233. Let a man guard himself against irritability in thought; let him be controlled in mind. Abandoning mental misconduct, let him practice good conduct in thought.

234. The wise are controlled in bodily action, controlled in speech and controlled in thought. They are truly well-controlled.
