\newpage
\chapter{Nirayavagga: Hell}
306. The liar goes to the state of woe; also he who, having done (wrong), says, "I did not do it." Men of base actions both, on departing they share the same destiny in the other world.

307. There are many evil characters and uncontrolled men wearing the saffron robe. These wicked men will be born in states of woe because of their evil deeds.

308. It would be better to swallow a red-hot iron ball, blazing like fire, than as an immoral and uncontrolled monk to eat the alms of the people.

309. Four misfortunes befall the reckless man who consorts with another's wife: acquisition of demerit, disturbed sleep, ill-repute, and (rebirth in) states of woe.

310. Such a man acquires demerit and an unhappy birth in the future. Brief is the pleasure of the frightened man and woman, and the king imposes heavy punishment. Hence, let no man consort with another's wife.

311. Just as kusa grass wrongly handled cuts the hand, even so, a recluse's life wrongly lived drags one to states of woe.

312. Any loose act, any corrupt observance, any life of questionable celibacy — none of these bear much fruit.

313. If anything is to be done, let one do it with sustained vigor. A lax monastic life stirs up the dust of passions all the more.

314. An evil deed is better left undone, for such a deed torments one afterwards. But a good deed is better done, doing which one repents not later.

315. Just as a border city is closely guarded both within and without, even so, guard yourself. Do not let slip this opportunity (for spiritual growth). For those who let slip this opportunity grieve indeed when consigned to hell.

316. Those who are ashamed of what they should not be ashamed of, and are not ashamed of what they should be ashamed of — upholding false views, they go to states of woe.

317. Those who see something to fear where there is nothing to fear, and see nothing to fear where there is something to fear — upholding false views, they go to states of woe.

318. Those who imagine evil where there is none, and do not see evil where it is — upholding false views, they go to states of woe.

319. Those who discern the wrong as wrong and the right as right — upholding right views, they go to realms of bliss.
