\newpage
\chapter{Arahantavagga: The Arahant or Perfected One}

90. The fever of passion exists not for him who has completed the journey, who is sorrowless and wholly set free, and has broken all ties.

91. The mindful ones exert themselves. They are not attached to any home; like swans that abandon the lake, they leave home after home behind.

92. Those who do not accumulate and are wise regarding food, whose object is the Void, the Unconditioned Freedom — their track cannot be traced, like that of birds in the air.

93. He whose cankers are destroyed and who is not attached to food, whose object is the Void, the Unconditioned Freedom — his path cannot be traced, like that of birds in the air.

94. Even the gods hold dear the wise one, whose senses are subdued like horses well trained by a charioteer, whose pride is destroyed and who is free from the cankers.

95. There is no more worldly existence for the wise one who, like the earth, resents nothing, who is firm as a high pillar and as pure as a deep pool free from mud.

96. Calm is his thought, calm his speech, and calm his deed, who, truly knowing, is wholly freed, perfectly tranquil and wise.

97. The man who is without blind faith, who knows the Uncreated, who has severed all links, destroyed all causes (for karma, good and evil), and thrown out all desires — he, truly, is the most excellent of men.\footnote{In the Pali this verse presents a series of puns, and if the "underside" of each pun were to be translated, the verse would read thus: "The man who is faithless, ungrateful, a burglar, who destroys opportunities and eats vomit — he truly is the most excellent of men."}

98. Inspiring, indeed, is that place where Arahants dwell, be it a village, a forest, a vale, or a hill.

99. Inspiring are the forests in which worldlings find no pleasure. There the passionless will rejoice, for they seek no sensual pleasures.
